---
title: "Sending Emails with blastula"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

This project demonstrates how to use [**blastula**](https://github.com/rstudio/blastula) to send personalized emails with attachments to a list of clients stored in an Excel spreadsheet.


## Project structure

```
blastula-example/
├── assets/
│   └── pdf-example.pdf   # Attachment to send
├── clients.xlsx          # Client contact list
├── README.Rmd            # This file (renders to README.md)
└── send-emails.R         # Email sending script
```


## Setup

Install the required packages:

```{r}
install.packages(c("blastula", "readxl", "glue"))
```


## Client list

`clients.xlsx` contains the contact list with columns `email`, `first`, and `last`:

```{r}
library("readxl")

(clients <- read_xlsx("clients.xlsx"))
```



## Composing and sending emails

### Step 1: Store your SMTP credentials

Before sending, store your SMTP credentials in your system keyring. You only need to do this once:

```{r}
if (!requireNamespace("keyring")) install.packages("keyring")
library("blastula")

create_smtp_creds_key(
  id = "baylor",
  user = "david_kahle@baylor.edu",
  host = "smtp.office365.com",
  port = 587,
  use_ssl = TRUE
)
```

This will prompt you for your password. If your account has MFA enabled, you may need to generate an App Password through your Microsoft account security settings.

You can verify your stored credentials with:

```{r}
view_credential_keys()
```


### Step 2: Compose and send a personalized email to each client

```{r}
library("blastula")
library("readxl")
library("glue")


# read in the excel spreadsheet containing the email addresses
clients <- read_xlsx("clients.xlsx")


# loop over rows in excel spreadsheet, making emails, attach, and send
for (i in seq_len(nrow(clients))) {

  client <- clients[i, ]

  # make the email
  email <- compose_email(
    body = md(glue(
      "Hi {client$first},

      Please find the attached document for your review.

      Best,

      David"
    ))
  )

  # add attachments
  email <- add_attachment(email, file = "assets/pdf-example.pdf")

  # send the email
  smtp_send(
    email,
    from = "david_kahle@baylor.edu",
    to = client$email,
    subject = glue("Document for {client$first} {client$last}"),
    credentials = creds_key("baylor")
  )

  message(glue("Sent to {client$first} {client$last} ({client$email})"))
}
```


## Standalone script

The same workflow is available as a standalone script in `send-emails.R`, ready to run from RStudio or the command line:

```bash
Rscript send-emails.R
```
